The Noisy Work of Uncertainty Visualisation Research: A Review

### 2024-05-14 paper.qmd
two small edits

Visualisation authors are almost unanimous in commenting that the "information" in two plots must be the same in order for the visual techniques to be compared [@Cleveland1984; @Kinkeldey2014]. @Kinkeldey2014 adopts an existing definition also that suggests two graphics are informationally equivalent if all the information in one plot is inferable from the other and vice-versa, but adds that two plots are computationally equivalent if that information can be extracted from both plots with similar easy and speed. Many visualisation authors ignore this concept and simply compare two visualisations because they both belong to the class of "uncertainty visualisations". @Ibrekk1987 compared a 6 visualisations of a PDF,  a box plot, a CDF and a mean with a confidence interval on this basis. They found that people are better at extracting the mean from a plot when they are shown a plot that contains a mean with a confidence interval than when they are shown a PDF, a box plot, or a CDF (where the mean could not be read off the plot). @Hullman2015 compared the static error bars and violin plots of the marginal distributions of two variables ($A$ and $B$) to an animated plot that depicted outcomes of the joint distribution of $A$ and $B$ in each frame and found that the visualisation of the joint distibution was better at answering questions about the joint distribution than the visualisations of the marginal distribution. @Hofman2020 commented that "theoretically" the sampling distribution of the mean and the prediction interval of a new observation are equal "so long as one knows the sample size", but does not seem to provide participants with that sample size, or recognise the assumptions and background knowledge that would be required to compare the two. @Hofman2020 and @Zhang2022 compared prediction and sampling distributions because they are both "uncertainty" that is typically depicted around the mean. They found that people are better at answering questions about a prediction interval when shown a prediction interval instead of a sampling distribution.

@@

**The following sections are in the middle of being reorganized and rewritten so it will not flow on from the previous sections**

>>>>

Visualisation authors are almost unanimous in commenting that the "information" in two plots must be the same in order for the visual techniques to be compared [@Cleveland1984; @Kinkeldey2014]. @Kinkeldey2014 adopts an existing definition also that suggests two graphics are informationally equivalent if all the information in one plot is inferable from the other and vice-versa, but adds that two plots are computationally equivalent if that information can be extracted from both plots with similar easy and speed. Many visualisation authors ignore this concept and simply compare two visualisations because they both belong to the class of "uncertainty visualisations". @Ibrekk1987 compared a 6 visualisations of a PDF,  a box plot, a CDF and a mean with a confidence interval on this basis. They found that people are better at extracting the mean from a plot when they are shown a plot that contains a mean with a confidence interval than when they are shown a box plot, or other visualisations that did not allow for the mean to just be read off the plot. @Hullman2015 compared the static error bars and violin plots of the marginal distributions of two variables ($A$ and $B$) to an animated plot that depicted outcomes of the joint distribution of $A$ and $B$ in each frame and found that the visualisation of the joint distibution was better at answering questions about the joint distribution than the visualisations of the marginal distribution. @Hofman2020 commented that "theoretically" the sampling distribution of the mean and the prediction interval of a new observation are equal "so long as one knows the sample size", but does not seem to provide participants with that sample size, or recognise the assumptions and background knowledge that would be required to compare the two. @Hofman2020 and @Zhang2022 compared prediction and sampling distributions because they are both "uncertainty" that is typically depicted around the mean. They found that people are better at answering questions about a prediction interval when shown a prediction interval instead of a sampling distribution.

@@

**UP TO HERE FOR RESTRUCTURE/REWRITE**
